Comparison of Terrain Following and Cut Cell Grids using a Non-Hydrostatic Model

### 2015-05-15 msc-publication.tex
update table of advection stats, discuss domain differences and interp of z for calculating psi

\TODO{take care to explain domain and grid because of our previous differences with Schaer's results} The domain is \SI{300}{\kilo\meter} wide and \SI{25}{\kilo\meter} high, discretized onto a grid with \(\Delta x = \SI{1}{\kilo\meter}\) and \(\Delta z^\star = \SI{500}{\meter}\).  The terrain is wave-shaped, specified by the surface height, \(h\), such that
\begin{subequations}
\begin{align}
   h(x) &= h^\star \cos^2 ( \alpha x )
%
\intertext{where}
%
   h^\star(x) &= \left\{ \begin{array}{l l}
       h_0 \cos^2 ( \beta x ) & \quad \text{if $| x | < a$} \\
	0 & \quad \text{otherwise}
    \end{array} \right.
\end{align}
\end{subequations}
where $a = \SI{25}{\kilo\meter}$ is the mountain envelope half-width, $h_0 = \SI{3}{\kilo\meter}$ is the maximum mountain height, $\lambda = \SI{8}{\kilo\meter}$ is the wavelength, \(\alpha = \pi / \lambda\) and \(\beta = \pi / 2a\).  On the SLEVE grid, the large-scale component $h_1$ is given by \(h_1(x) = h^\star(x) / 2\)
and $s_1 = \SI{15}{\kilo\meter}$ is the large scale height, and $s_2 = \SI{2.5}{\kilo\meter}$ is the small scale height.  The optimisation of SLEVE by \citet{leuenberger2010} is not used, so the exponent $n = 1$.

@@

By \(t = \SI{10000}{\second}\), the tracer suffers from diffusion and distortion on the BTF grid using the centred linear scheme.  Some artefacts remain above the mountain peak, and a computational mode is seen as a grid-scale oscillation that travels in the opposite direction to the wind.

@@

For the cubic upwind-biased scheme, error is largest on the BTF grid with \(\ell^2 = \num{0.00767}\) but significant reduced on the SLEVE grid with \(\ell^2 = \num{0.00108}\).  The error is approximately halved by changing from the SLEVE grid to the cut cell grid.

@@

\begin{table*}[t]
	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.}
\label{tab:advection}
%
\centering
\footnotesize
\begin{tabular}{ r @{\hspace{1.0em}} l @{\hspace{1.0em}} l @{\hspace{1.0em}} l | l @{\hspace{1.0em}} l @{\hspace{1.0em}} l | l @{\hspace{1.0em}} l @{\hspace{1.0em}} l | l @{\hspace{1.0em}} l @{\hspace{1.0em}} l}
\hline\hline
& \multicolumn{6}{c}{Horizontal advection} & \multicolumn{6}{c}{Terrain following advection} \\
& \multicolumn{3}{c}{Centred linear} & \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{3}{c}{Centred linear} & \multicolumn{3}{c}{Cubic upwind-biased} \\
& \(\ell^2\) error & min & max & \(\ell^2\) error & min & max & \(\ell^2\) error & min & max & \(\ell^2\) error & min & max \\
\hline
Analytic  & 0 & 0 & 1 & 0 & 0 & 1 & 0 & 0 & 1 & 0 & 0 & 1 \\
BTF 	  & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{9999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} \\
SLEVE & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{9999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} \\
Cut cell & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{9999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} \\
No terrain & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & {---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
\hline
\end{tabular}
\end{table*}

>>>>

The domain is \SI{301}{\kilo\meter} wide and \SI{25}{\kilo\meter} high, discretized onto a grid with \(\Delta x = \SI{1}{\kilo\meter}\) and \(\Delta z^\star = \SI{500}{\meter}\).  The domain specified by \citet{schaer2002} is \SI{300}{\kilo\meter} between the outermost cell centres.  In order to reproduce the result from \citet{schaer2002}, the domain has been extended horizontally by \(\Delta x/2\) \si{\meter} in both directions so that the distance between the outermost cell centres is still \SI{300}{\kilo\meter}.

The terrain is wave-shaped, specified by the surface height, \(h\), such that
\begin{subequations}
\begin{align}
   h(x) &= h^\star \cos^2 ( \alpha x )
%
\intertext{where}
%
   h^\star(x) &= \left\{ \begin{array}{l l}
       h_0 \cos^2 ( \beta x ) & \quad \text{if $| x | < a$} \\
	0 & \quad \text{otherwise}
    \end{array} \right.
\end{align}
\end{subequations}
where $a = \SI{25}{\kilo\meter}$ is the mountain envelope half-width, $h_0 = \SI{3}{\kilo\meter}$ is the maximum mountain height, $\lambda = \SI{8}{\kilo\meter}$ is the wavelength, \(\alpha = \pi / \lambda\) and \(\beta = \pi / 2a\).  On the SLEVE grid, the large-scale component $h_1$ is given by \(h_1(x) = h^\star(x) / 2\)
and $s_1 = \SI{15}{\kilo\meter}$ is the large scale height, and $s_2 = \SI{2.5}{\kilo\meter}$ is the small scale height.  The optimisation of SLEVE by \citet{leuenberger2010} is not used, so the exponent $n = 1$.

@@

By \(t = \SI{10000}{\second}\), the tracer suffers from diffusion and distortion on the BTF grid using the centred linear scheme.  Some artefacts remain above the mountain peak, and the computational mode is seen as a grid-scale oscillation that travels in the opposite direction to the wind.

@@

The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002}.  Error is largest on the BTF grid with \(\ell^2 = \num{0.00791}\) but significant reduced on the SLEVE grid with \(\ell^2 = \num{0.00108}\).  The error is approximately halved by changing from the SLEVE grid to the cut cell grid.

Tracer minima and maxima are lower than values presented by \citet{schaer2002} because we do not interpolate the geometric height of cell faces from adjacent cell centres in order to calculate the streamfunction, \(\Psi(z)\).  Removing the interpolation from Sch\"{a}r's code results in minima and maxima that are equal to those obtained from the OpenFOAM advection solver presented here.

The centred linear scheme is sensitive to changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell^2$ error on the BTF grid rises from \num{0.0210} to \num{0.0318}.  The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell^2$ error decreasing from \num{0.00791} to \num{0.00767}.

@@

\begin{table*}[t]
	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.}
\label{tab:advection}
%
\centering
\footnotesize
\begin{tabular}{ r @{\hspace{1.0em}} l @{\hspace{1.0em}} l @{\hspace{1.0em}} l | l @{\hspace{1.0em}} l @{\hspace{1.0em}} l | l @{\hspace{1.0em}} l @{\hspace{1.0em}} l | l @{\hspace{1.0em}} l @{\hspace{1.0em}} l}
\hline\hline
& \multicolumn{6}{c}{Horizontal advection} & \multicolumn{6}{c}{Terrain following advection} \\
& \multicolumn{3}{c}{Centred linear} & \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{3}{c}{Centred linear} & \multicolumn{3}{c}{Cubic upwind-biased} \\
& \(\ell^2\) error & min & max & \(\ell^2\) error & min & max & \(\ell^2\) error & min & max & \(\ell^2\) error & min & max \\
\hline
Analytic   & 0 & 0 & 1 & 0 & 0 & 1 & 0 & 0 & 1 & 0 & 0 & 1 \\
% 	     l2		     min	     max	   l2		    min		     max           l2                        min                       max          l2                         min                       max
BTF 	   & \num{0.0210}  & \num{-0.275}  & \num{0.925} & \num{0.00791}  & \num{-0.0446}  & \num{0.925} & \num{0.00251}           & \num{-0.0245}           & \num{0.985} & \num{0.00154}           & \num{-0.0110}           & \num{0.983} \\
SLEVE      & \num{0.00233} & \num{-0.0252} & \num{0.985} & \num{0.00108}  & \num{-0.0106}  & \num{0.982} & \num{0.0173}            & \num{-0.120}            & \num{0.950} & \num{0.0120}            & \num{-0.0263}           & \num{0.865} \\
Cut cell   & \num{0.00224} & \num{-0.0251} & \num{0.985} & \num{0.000577} & \num{-0.00674} & \num{0.983} & \num{13.6}              & \num{-623}              & \num{3480}  & \num{0.0134}            & \num{-0.028}            & \num{0.851} \\
No terrain & \num{0.00223} & \num{-0.0251} & \num{0.985} & \num{0.000576} & \num{-0.00674} & \num{0.983} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & {---}       & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
\hline
\end{tabular}
\end{table*}
